---
title: "Pricing FInancial Derivatives"
subtitle: "Project 2"
author: "Angus McKay, Laura Roman, Euan Dowers, Veronika Kyuchukova"
date: "February 20, 2017"
output: pdf_document
---

##Exercise 2

We are given a Stochastic Differential Equation:

\begin{align}
dX_{t}=a(m-X_{t})dt + \sigma dB_{t}\\
X_{0}=x
\end{align}
where a>0, $\sigma$ >0, m$\ge$ 0

\begin{align}
f(X_{t},t)= e^{at}X_{t}\\
\Rightarrow^{From~Itô}~ f(X_{t},t)= a X_{t} e^{at}dt + e^{at}dX_{t}
\end{align}

Substituting $dX_{t}$ with (1) we get:

\begin{align}
f(X_{t},t) &= a X_{t} e^{at} dt + a e^{at}mdt -ae^{at}X_{t}dt + e^{at}\sigma dB_{t}\\
\Rightarrow &= a m e^{at} dt + e^{at}\sigma dB_{t}
\end{align}

Now integrating over $[0,t]$:
\begin{align}
f(X_{t},t) = e^{at}X_{t}&= X_{0}+ \int_{0}^{t} a m e^{as} ds + \int_{0}^{t} e^{as}\sigma dB_{s}\\
\Rightarrow X_{t}&= xe^{-at}+ e^{-at}[me^{as}]_{0}^{t}+ e^{-at}\sigma \int_{0}^{t} e^{as}dB_{s}\\
X_{t}&= xe^{-at}+m- me^{at}+ e^{-at}\sigma \int_{0}^{t} e^{as}dB_{s}\\
X_{t}&=m+(x-m)e^{-at}+\sigma e^{-at}\int_{0}^{t}e^{as}dB_{s}
\end{align}

\begin{align}
E[X_{t}]= xe^{-at}+m(1-e^{-at})
\end{align}


\begin{align}
Var[X_{t}] &= E[(X_{t}-E[X_{t}])^2]\\
           &= E\left[ \int_{0}^{t} \sigma^{2} e^{2(s-t)a}dB_{s}\right]\\
           &= \sigma^{2}e^{-2at}E\left[ \int_{0}^{t}  e^{2sa}dB_{s}\right]\\
           &= \sigma^{2}e^{-2at} \left[ \frac{e^{2sa}}{2a} \right]_{0}^{t} \\
           &= \frac{\sigma^{2}}{2a} e^{-at}( e^{at}-1)\\
           &= \frac{\sigma^2}{2a}(1-e^{2at})
\end{align}



As $t\rightarrow \infty,~xe^{-at}\rightarrow0,~(1-e^{-at})\rightarrow1$ 

\begin{align}
\Rightarrow lim_{t \rightarrow \infty}E[X_{t}]= m
\end{align}



```{r, eval=TRUE, echo=FALSE}
library(ggplot2)

set.seed(444)

N <- 1000
T <- 1
x <- 0.7
dt <- 1 / N

Vasicek_function <- function(m, sigma, a){
  dX <- numeric ( N +1)
  dX[1] <- x
  Z <- rnorm ( N, 0, 1 )
  for ( i in 1: N ){
    dX [i +1] <- dX [ i ] + a*(m -  dX [ i ]) * dt + sigma * sqrt ( dt ) * Z [ i ]
  dX <- ts (dX , start =0 , deltat =1 / N )
  }
  return(dX)
}

V<-matrix(0,7,N+1)
m<-c(2,1,5,2,2,2,2)
sigma<-c(2,2,2,5,1,2,2)
a<-c(2,2,2,2,2,5,1)

for (i in 1:nrow(V)){
  V[i,]<-Vasicek_function(m[i],sigma[i],a[i])
}

color<-c("red","blue","yellow","black","green","purple","pink")

plot(V[1,], main= "Trajectories of Vasicek model",
     col="red", ylim=c(-3,6.5), xlab="t", ylab="X",type = "l")
grid(5,5,col="black")

for (i in 2:nrow(V)){
lines(V[i,], col= color[i])
}

legend("topleft",
       c("m=2, sigma=2, a=2","m=1, sigma=2, a=2","m=5, sigma=2, a=2",
         "m=2, sigma=5, a=2","m=2, sigma=1, a=2","m=2, sigma=2, a=5","m=2, sigma=2, a=1"),
       col=c("red","blue","yellow","black",
               "green","purple","pink"),
       lty=c(1,1),
       cex=0.58)

```














\newpage
##Appendix
```{r, eval=FALSE}
library(ggplot2)


N <- 1000
T <- 1
x <- 0.7
dt <- 1 / N

Vasicek_function <- function(m, sigma, a){
  dX <- numeric ( N +1)
  dX[1] <- x
  Z <- rnorm ( N, 0, 1 )
  for ( i in 1: N ){
    dX [i +1] <- dX [ i ] + a*(m -  dX [ i ]) * dt + sigma * sqrt ( dt ) * Z [ i ]
  dX <- ts (dX , start =0 , deltat =1 / N )
  return(dX)
  }
}

V<-matrix(0,7,N+1)
m<-c(2,1,5,2,2,2,2)
sigma<-c(2,2,2,5,1,2,2)
a<-c(2,2,2,2,2,5,1)

for (i in 1:nrow(V)){
  V[i,]<-Vasicek_function(m[i],sigma[i],a[i])
}
V
color<-c("red","blue","yellow","black","green","purple","pink")

plot(V[1,], main= "Trajectories of Vasicek model",
     col="red", ylim=c(-3,6.5), xlab="t", ylab="X",type = "l")
grid(5,5,col="black")

for (i in 2:nrow(V)){
lines(V[i,], col= color[i])
}

legend("topleft",
       c("m=2, sigma=2, a=2","m=1, sigma=2, a=2","m=5, sigma=2, a=2",
         "m=2, sigma=5, a=2","m=2, sigma=1, a=2","m=2, sigma=2, a=5","m=2, sigma=2, a=1"),
       col=c("red","blue","yellow","black",
               "green","purple","pink"),
       lty=c(1,1),
       cex=0.58)

```
