---
title: "Pricing Financial Derivatives Project 2"
author: "Angus McKay, Laura Roman, Euan Dowers, Veronika Kyuchukova"
date: "February 2, 2017"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Exercise 1: Black-Scholes model
Consider the Black-Scholes model:

$$dX_t=2X_tdt+1.5X_tdB_t\text{, with }X_0=1$$
The solution to this model is:

$$X_t=e^{1.5B_t+(2-1.5^2/2)t}$$

```{r Exercise 1, echo=FALSE}
library(ggplot2)

Sn <- function(n,T = 5){
  # Create function that produces random walk with step size sqrt(T/n)
  stepsize <- sqrt(T/n)
  Z <- rbinom(n+1,1,0.5)
  Z[Z==1] <- stepsize
  Z[Z==0] <- -stepsize
  B <- rep(0,n+1)
  for ( i in 1:n+1) { 
    B[i] = B[i-1] + Z[i] }
  return(B)
}

# Now plot the process Xn(t) for n = 1000 using a Brownian path
T <- 5
n <- 1000
B <- Sn(n)
t <- seq(0,T,length.out = length(B))
X <- exp(1.5*B+(2-1.5**2/2)*t)
dd1 <- as.data.frame(cbind(t,X))
q1 <- ggplot(data=dd1, aes(x=t)) + 
  geom_line(aes(y=X)) +
  ggtitle("projection of Xt using brownian path")+
  ylab("Xt") + xlab("t") 
q1

# Using the same Brownian with Euler's scheme
E <- rep(1,n+1)
for ( i in 1:n+1) { 
    E[i] = E[i-1] + 1.5*E[i-1]*(B[i]-B[i-1]) + 2*E[i-1]*T/n }
dd2 <- as.data.frame(cbind(t,E))
q2 <- ggplot(data=dd2, aes(x=t)) + 
  geom_line(aes(y=E)) +
  ggtitle("projection of Xt using Euler's scheme")+
  ylab("Xt") + xlab("t") 
q2

```

As expected the results produce very similar outputs. When looking at the figures used in the charts the numbers are actually slightly different but it is not noticeable in the charts.

## Exercise 2: Vasicek model for interest rates

```{r Exercise 2, echo=FALSE}
m1 <- 2
sigma1 <- 2
a1 <- 2
x1 <- 2/3
V1 <- rep(x1,n+1)
for ( i in 1:n+1) { 
    V1[i] = V1[i-1] + sigma1*(B[i]-B[i-1]) + a1*(m1-V1[i-1])*T/n }

m2 <- 1
V2 <- rep(x1,n+1)
for ( i in 1:n+1) { 
    V2[i] = V2[i-1] + sigma1*(B[i]-B[i-1]) + a1*(m2-V2[i-1])*T/n }

m3 <- 5
V3 <- rep(x1,n+1)
for ( i in 1:n+1) { 
    V3[i] = V3[i-1] + sigma1*(B[i]-B[i-1]) + a1*(m3-V3[i-1])*T/n }

sigma4 <- 5
V4 <- rep(x1,n+1)
for ( i in 1:n+1) { 
    V4[i] = V4[i-1] + sigma4*(B[i]-B[i-1]) + a1*(m1-V4[i-1])*T/n }

sigma5 <- 1
V5 <- rep(x1,n+1)
for ( i in 1:n+1) { 
    V5[i] = V5[i-1] + sigma5*(B[i]-B[i-1]) + a1*(m1-V5[i-1])*T/n }

a6 <- 5
V6 <- rep(x1,n+1)
for ( i in 1:n+1) { 
    V6[i] = V6[i-1] + sigma1*(B[i]-B[i-1]) + a6*(m1-V6[i-1])*T/n }

a7 <- 1
V7 <- rep(x1,n+1)
for ( i in 1:n+1) { 
    V7[i] = V7[i-1] + sigma1*(B[i]-B[i-1]) + a7*(m1-V7[i-1])*T/n }

ggplot() +
  geom_line(aes(x = t, y = V1, colour = "m=2, sigma=2, a=2"), size = 0.2) +
  geom_line(aes(x = t, y = V2, colour = "m=1, sigma=2, a=2"), size = 0.2) +
  geom_line(aes(x = t, y = V3, colour = "m=5, sigma=2, a=2"), size = 0.2) +
  geom_line(aes(x = t, y = V4, colour = "m=2, sigma=5, a=2"), size = 0.2) +
  geom_line(aes(x = t, y = V5, colour = "m=2, sigma=1, a=2"), size = 0.2) +
  geom_line(aes(x = t, y = V6, colour = "m=2, sigma=2, a=5"), size = 0.2) +
  geom_line(aes(x = t, y = V7, colour = "m=2, sigma=2, a=1"), size = 0.2) +
  labs(title = "Trajectories of Vasicek model", x = "t", y = "X")


```

The mean reverting phenomenon can be observed, particularly when comparing the trajectories where m = 2 and m = 5. When sigma is higher it can be seen that the mean reversion is exhibited but there is much more volatility around the mean value.

## Exercise 3: Vasicek model for interest rates

```{r Exercise 3, echo=FALSE}
alpha1 <- 6
beta1 <- 1
sig1 <- 1
r0 <- 0
CIR1 <- rep(r0,n+1)
for ( i in 1:n+1) { 
    CIR1[i] = CIR1[i-1] + sig1*sqrt(CIR1[i-1])*(B[i]-B[i-1]) + (alpha1-beta1*CIR1[i-1])*T/n }

beta2 <- 2
sig2 <- 2
CIR2 <- rep(r0,n+1)
for ( i in 1:n+1) { 
    CIR2[i] = CIR2[i-1] + sig2*sqrt(CIR2[i-1])*(B[i]-B[i-1]) + (alpha1-beta2*CIR2[i-1])*T/n }

beta3 <- 5
sig3 <- 3
CIR3 <- rep(r0,n+1)
for ( i in 1:n+1) { 
    CIR3[i] = CIR3[i-1] + sig3*sqrt(CIR3[i-1])*(B[i]-B[i-1]) + (alpha1-beta3*CIR3[i-1])*T/n }

alpha4 <- 2
beta4 <- 2
sig4 <- 2
CIR4 <- rep(r0,n+1)
for ( i in 1:n+1) { 
    CIR4[i] = CIR4[i-1] + sig4*sqrt(CIR4[i-1])*(B[i]-B[i-1]) + (alpha4-beta4*CIR4[i-1])*T/n }

alpha5 <- 1
beta5 <- 1
sig5 <- 1
CIR5 <- rep(r0,n+1)
for ( i in 1:n+1) { 
    CIR5[i] = CIR5[i-1] + sig5*sqrt(CIR5[i-1])*(B[i]-B[i-1]) + (alpha5-beta5*CIR5[i-1])*T/n }

alpha6 <- 1
beta6 <- 2
sig6 <- 1
CIR6 <- rep(r0,n+1)
for ( i in 1:n+1) { 
    CIR6[i] = CIR6[i-1] + sig6*sqrt(CIR6[i-1])*(B[i]-B[i-1]) + (alpha6-beta6*CIR6[i-1])*T/n }

alpha7 <- 1
beta7 <- 0.1
sig7 <- 3.7
CIR7 <- rep(r0,n+1)
for ( i in 1:n+1) { 
    CIR7[i] = CIR7[i-1] + sig7*sqrt(CIR7[i-1])*(B[i]-B[i-1]) + (alpha7-beta7*CIR7[i-1])*T/n }

alpha8 <- 1
beta8 <- 10
sig8 <- 1
CIR8 <- rep(r0,n+1)
for ( i in 1:n+1) { 
    CIR8[i] = CIR8[i-1] + sig8*sqrt(CIR8[i-1])*(B[i]-B[i-1]) + (alpha8-beta8*CIR8[i-1])*T/n }

ggplot() +
  geom_line(aes(x = t, y = CIR1, colour = 'alpha=6, beta=1, sigma=1'), size = 0.2) +
  geom_line(aes(x = t, y = CIR2, colour = 'alpha=6, beta=2, sigma=2'), size = 0.2) +
  geom_line(aes(x = t, y = CIR3, colour = 'alpha=6, beta=5, sigma=3'), size = 0.2) +
  geom_line(aes(x = t, y = CIR4, colour = 'alpha=2, beta=2, sigma=2'), size = 0.2) +
  geom_line(aes(x = t, y = CIR5, colour = 'alpha=1, beta=1, sigma=1'), size = 0.2) +
  geom_line(aes(x = t, y = CIR6, colour = 'alpha=1, beta=2, sigma=1'), size = 0.2) +
  geom_line(aes(x = t, y = CIR8, colour = 'alpha=1, beta=10, sigma=1'), size = 0.2) +
  labs(title = "Trajectories of CIR model", x = "t", y = "r")



```



## Appendix
##### R code
```{r appendix 1, ref.label='Exercise 1', eval = FALSE}
```

```{r appendix 1, ref.label='Exercise 2', eval = FALSE}
```

```{r appendix 1, ref.label='Exercise 3', eval = FALSE}
```
